CYO Project Report - Delivery Estimation System

### 2021-07-17 finalReportPrep.Rmd
stabalized report with models and did cleanup

Delivery Performance:

You will also be able to work through delivery performance and find ways to optimize delivery times.

Attention

- An order might have multiple items.
- Each item might be fulfilled by a distinct seller.
- All text identifying stores and partners where replaced by the names of Game of Thrones great houses.

olist_customers_dataset.csv
This dataset has information about the customer and its location. Use it to identify unique customers in the orders dataset and to find the orders delivery location.
At our system each order is assigned to a unique customerid. This means that the same customer will get different ids for different orders. The purpose of having a customerunique_id on the dataset is to allow you to identify customers that made repurchases at the store. Otherwise you would find that each order had a different customer associated with.

olist_order_reviews_dataset.csv
This dataset includes data about the reviews made by the customers.

After a customer purchases the product from Olist Store a seller gets notified to fulfill that order. Once the customer receives the product, or the estimated delivery date is due, the customer gets a satisfaction survey by email where he can give a note for the purchase experience and write down some comments.

olist_orders_dataset.csv
This is the core dataset. From each order you might find all other information.

olist_products_dataset.csv
Products Dataset
This dataset includes data about the products sold by Olist.

@@

```{r RMSE, echo=FALSE}
# Performance metric
RMSE <- function(true_outcomes, predicted_outcomes) {
  sqrt(mean((true_outcomes - predicted_outcomes)^2))
}
```

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>%
  semi_join(train_subset,by='seller_id') %>% 
  semi_join(train_subset,by="customer_zip_code_prefix") %>% 
  semi_join(train_subset,by="customer_loc")

penalty_data <- train_subset %>% get_delay_in_handover_to_carrier_per_seller_id()

delay_in_payment_approval_data <- train_subset %>% get_delay_in_payment_approval_per_customer_zip_code()

@@

```{r Prepare training and test data for ,include=FALSE}
# penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
train_subset <- train_subset %>% 
  left_join(penalty_data,by='seller_id') %>% 
  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)

test_subset <- test_subset %>% 
  left_join(penalty_data,by='seller_id') %>% 
  left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix') %>% 
  select(-handed_over_to_carrier_date, -shipping_limit_date,-order_purchase_timestamp,-payment_approved_at)
```

@@

@@

@@

```{r Prepare data set for training, include = FALSE}
# prepares data for training and testing
# returns list of data frames. First list has training dataframe. Second list has test data frame
prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
  print('Adding column for delay in handover')
  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()

  #we will use penalty_data computed using training data in test data.
  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
  if (cleanup) {
    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
  }

  print('Adding column for delay in payment approval')
  delay_in_payment_approval <- train_ds %>% get_delay_in_payment_approval_per_customer_zip_code()

  #we will use delay_in_payment_approval computed using training data in test data.
  train_ds <- train_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
  test_ds <- test_ds %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')
  if (cleanup) {
    train_df <- train_df %>% select(-customer_zip_code_prefix)
    test_df <- test_df %>% select(-customer_zip_code_prefix)
  }

@@

```{r Apply Target encoding for product_category_name, echo = FALSE}
#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.

  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-seller_id,-customer_loc, -customer_zip_code_prefix, -customer_state)

@@

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name')

  weight_rmses <- sapply(weights, function(weight) {
      product_category_effects <- train_weight_set %>% 
        group_by(product_category_name) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

@@

```{r Post regularization for product_category effect, include=FALSE}
weight_to_use <- target_encoding_weights %>% filter(effect=='product_category') %>% pull(weight)

product_category_effects <- prepare_product_category_effect(train_subset,weight_to_use)
dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_subset,test_subset,TRUE)

train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]
```

```{r Apply Target encoding for seller_id, echo = FALSE}
#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.

  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-customer_loc, -customer_zip_code_prefix, -customer_state)

@@

  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='seller_id')

  weight_rmses <- sapply(weights, function(weight) {
      seller_id_effects <- train_weight_set %>% 
        group_by(seller_id) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

weight_to_use <- mean(weight_to_use_per_partition)  
# 3

```{r Post regularization for seller_id effect, include=FALSE}
weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)

seller_id_effects <- prepare_seller_id_effect(train_subset,weight_to_use)
dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_subset,test_subset,TRUE)

train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]
```

```{r Apply Target encoding for customer_state, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc,-customer_zip_code_prefix)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

  weight_rmses <- sapply(weights, function(weight) {
      customer_state_effects <- train_weight_set %>% 
        group_by(customer_state) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)

@@

@@

```{r Post regularization with consumer state effect, include=FALSE}
weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)

customer_state_effects <- prepare_customer_state_effect(train_subset,weight_to_use)
dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_subset,test_subset,TRUE)

train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]
```

```{r Apply Target encoding for customer_loc, echo = FALSE}
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_zip_code_prefix)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

  test_weight_set <- test_weight_set %>% semi_join(train_weight_set,by='customer_loc')

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(20,27,34)),trControl=tuningControl)

@@

```{r Post regularization with customer loc effect, include=FALSE}
weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)

customer_loc_effects <- prepare_customer_loc_effect(train_subset,weight_to_use)
dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_subset,test_subset,TRUE)

train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]  
```

@@

```{r knn model data scaled under test, eval=FALSE, include=FALSE}
## TODO make echo=FALSE 
# TODO review and improve or remove this as required
set.seed(seed=1,sample.kind = 'Rounding')
preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))

@@

tuningControl=trainControl(method = "cv", number = 5, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)

@@

performance_metric <- performance_metric %>% 
  add_row(type='knn model',
          value=RMSE((scaled_test_subset %>% pull(number_of_hours)),predictions))

performance_metric
#183.8551 for k=70

@@

```{r rf model, eval = FALSE, include= FALSE}
# TODO make it echo=FALSE
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 4)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5)),trControl=tuningControl,ntree=200)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

performance_metric <- performance_metric %>% 
  add_row(type='random forest model',
          value=RMSE((test_subset %>% pull(number_of_hours)),predictions))

performance_metric
# 174.0326 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.

@@

>>>>

@@

```{r RMSE, include=FALSE}
# Performance metric
RMSE <- function(true_outcomes, predicted_outcomes) {
  sqrt(mean((true_outcomes - predicted_outcomes)^2))
}
```

@@

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]

# ensure test_subset has products only for concerned product categories and seller_ids which are in train_subset
test_subset <- test_subset %>% semi_join(train_subset,by='product_category_name') %>%
  semi_join(train_subset,by='seller_id') %>% 
  semi_join(train_subset,by="customer_zip_code_prefix") %>% 
  semi_join(train_subset,by="customer_loc")

@@

@@

```{r Prepare delay in handover to carried data, include = FALSE}
# prepare delay_in_handover_to_carrier_data based on regularisation weight we have identified
prepare_delay_in_handover_to_carrier <- function(train_df, weight_to_use) {
  global_mean <- mean(train_df %>% 
                        get_delay_in_handover_to_carrier_per_seller_id() %>% 
                        pull(mean_delay_in_handover))

  delay_in_handover_to_carrier_data_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

      delay_in_handover_to_carrier_data <- train_weight_set %>% 
        get_delay_in_handover_to_carrier_per_seller_id()

      delay_in_handover_to_carrier_data
    }
  )

  dhc <- delay_in_handover_to_carrier_data_list[[1]]
  for (index in 2:10) {
     dhc <- dhc %>% full_join(delay_in_handover_to_carrier_data_list[[index]],by='seller_id', suffix=c("1",index))
  }

  # wherever value is NA, we have used global mean
  dhc <- dhc %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))

  # as different partitions may not have exactly same data set combined the results
  delay_in_handover_to_carrier_data <- tibble(dhc[,1],
                                              mean_delay_in_handover=rowMeans(dhc[,-1],
                                                                              na.rm = TRUE))

  #delay_in_handover_to_carrier_data is ready
  delay_in_handover_to_carrier_data
}
```

```{r Prepare delay in payment approval, include = FALSE}
# prepare delay_in_payment_approval based on regularisation weight we have identified
prepare_delay_in_payment_approval <- function(train_df, weight_to_use) {
  global_mean <- mean(train_df %>% 
                        get_delay_in_payment_approval_per_customer_zip_code() %>% 
                        pull(mean_delay_in_payment_approval))

  delay_in_payment_approval_data_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

      delay_in_payment_approval_data <- train_weight_set %>% 
        get_delay_in_payment_approval_per_customer_zip_code()

      delay_in_payment_approval_data
    }
  )

  dpa <- delay_in_payment_approval_data_list[[1]]
  for (index in 2:10) {
     dpa <- dpa %>% full_join(delay_in_payment_approval_data_list[[index]],by='customer_zip_code_prefix', suffix=c("1",index))
  }

  # wherever value is NA, we have used global mean
  dpa <- dpa %>% mutate(across(-customer_zip_code_prefix,~if_else(is.na(.x),global_mean,.x)))

  # as different partitions may not have exactly same data set combined the results
  delay_in_payment_approval_data <- tibble(dpa[,1],
                                              mean_delay_in_payment_approval=rowMeans(dpa[,-1],
                                                                              na.rm = TRUE))

  #delay_in_payment_approval_data is ready
  delay_in_payment_approval_data
}
```

@@

```{r Enhance training and test dataset for delay in payment approval, include=FALSE}
enhance_datasets_using_delay_in_payment_approval <- function(delay_in_payment_approval_data,train_df,test_df,cleanup) {
  #we will use delay_in_payment_approval_data computed using training data in test data.  
  train_df <- train_df %>%
    left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')

  test_df <- test_df %>% left_join(delay_in_payment_approval_data,by='customer_zip_code_prefix')

  #cleanup
  if (cleanup) {
    train_df <- train_df %>%
      select(-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
    test_df <- test_df %>% 
      select(-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
  }

@@

```{r Enhance training and testing datasets with derived features, include =FALSE}
# uses training data to derive features and then enhances training and test set with those features
enhance_datasets_with_derived_features <- function(train_ds,test_ds,cleanup) {
  print('Adding column for delay in handover')

  delay_in_handover_to_carrier_data <- prepare_delay_in_handover_to_carrier(train_ds,0)

  dataset_list <- enhance_datasets_using_delay_in_handover_to_carrier(delay_in_handover_to_carrier_data,train_ds,test_ds,cleanup)
  train_ds <- dataset_list[[1]]
  test_ds <- dataset_list[[2]]

  print('Adding column for delay in payment approval')
  delay_in_payment_approval_data <- prepare_delay_in_payment_approval(train_ds,0)

  #we will use delay_in_payment_approval computed using training data in test data.
  dataset_list <- enhance_datasets_using_delay_in_payment_approval(delay_in_payment_approval_data,train_ds,test_ds,cleanup)

  dataset_list
}  
```

```{r Prepare data set for training, include = FALSE}
# prepares data for training and testing
# returns list of data frames. First list has training dataframe. Second list has test data frame
prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
  test_ds <- test_ds %>%
    semi_join(train_ds,by='product_category_name') %>%
    semi_join(train_ds,by='seller_id') %>% 
    semi_join(train_ds,by='customer_zip_code_prefix') %>% 
    semi_join(train_ds,by='customer_state') %>% 
    semi_join(train_ds,by='customer_loc')

  #adding derived features and target encoded categorical features
  print('Adding derived features')
  dataset_list <- enhance_datasets_with_derived_features(train_ds,test_ds,cleanup)
  train_ds <- dataset_list[[1]]
  test_ds <- dataset_list[[2]]

@@

```{r Apply Target encoding for product_category_name, include = FALSE, eval=FALSE}
#TODO marked eval=FALSE to help generate report quickly
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)

@@

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_id') %>%
    semi_join(train_weight_set,by='customer_zip_code_prefix')

  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,TRUE)
  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)#,-customer_zip_code_prefix)
  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)#,-customer_zip_code_prefix)

  weight_rmses <- sapply(weights, function(weight) {
      product_category_effects <- train_weight_set %>% 
        group_by(product_category_name) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n)

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

```{r Apply Target encoding for seller_id, include = FALSE, eval=FALSE}
#TODO marked eval=FALSE to help generate report quickly
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc, -customer_state)

@@

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_id') %>% 
    semi_join(train_weight_set,by='customer_zip_code_prefix')

  ## adding derived features and target encoded categorical features
  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
  train_weight_set <- dataset_list[[1]]
  test_weight_set <- dataset_list[[2]]

  weight_to_use <- target_encoding_weights %>% 
    filter(effect=='product_category') %>% pull(weight)

  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)

  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
  ##

  weight_rmses <- sapply(weights, function(weight) {
      seller_id_effects <- train_weight_set %>% 
        group_by(seller_id) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n)

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

```{r Apply Target encoding for customer_state, include = FALSE,eval=FALSE}
#TODO marked eval=FALSE to help generate report quickly
  train_subset_for_encoding <- train_subset[1:40000,] %>% select(-customer_loc)
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_id') %>% 
    semi_join(train_weight_set,by='customer_zip_code_prefix') %>% 
    semi_join(train_weight_set,by='customer_state')

  ## adding derived features and target encoded categorical features
  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
  train_weight_set <- dataset_list[[1]]
  test_weight_set <- dataset_list[[2]]

  weight_to_use <- target_encoding_weights %>% 
    filter(effect=='product_category') %>% pull(weight)

  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)

  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)

  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)

  seller_id_effects <- prepare_seller_id_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_weight_set,test_weight_set,FALSE)

  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)
  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)
  ##

  weight_rmses <- sapply(weights, function(weight) {
      customer_state_effects <- train_weight_set %>% 
        group_by(customer_state) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n)

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

weight_to_use <- mean(weight_to_use_per_partition) 
# 10

@@

```{r Apply Target encoding for customer_loc, include = FALSE,eval=FALSE}
#TODO marked eval=FALSE to help generate report quickly
  train_subset_for_encoding <- train_subset[1:40000,]
  global_mean <- mean(train_subset_for_encoding$number_of_hours)

@@

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name') %>%
    semi_join(train_weight_set,by='seller_id') %>% 
    semi_join(train_weight_set,by='customer_zip_code_prefix') %>% 
    semi_join(train_weight_set,by='customer_state') %>% 
    semi_join(train_weight_set,by='customer_loc')

  ## adding derived features and target encoded categorical features
  dataset_list <- enhance_datasets_with_derived_features(train_weight_set,test_weight_set,FALSE)
  train_weight_set <- dataset_list[[1]]
  test_weight_set <- dataset_list[[2]]

  weight_to_use <- target_encoding_weights %>% 
    filter(effect=='product_category') %>% pull(weight)

  product_category_effects <- prepare_product_category_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_weight_set,test_weight_set,FALSE)

  train_weight_set <- dataset_list[[1]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)
  test_weight_set <- dataset_list[[2]] %>% select(-order_purchase_timestamp,-handed_over_to_carrier_date,-shipping_limit_date,-customer_zip_code_prefix,-payment_approved_at,-product_category_name)

  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)

  seller_id_effects <- prepare_seller_id_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_weight_set,test_weight_set,FALSE)

  train_weight_set <- dataset_list[[1]] %>% select(-seller_id)
  test_weight_set <- dataset_list[[2]] %>% select(-seller_id)

  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)

  customer_state_effects <- prepare_customer_state_effect(train_weight_set,weight_to_use)
  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_weight_set,test_weight_set,TRUE)

  train_weight_set <- dataset_list[[1]]
  test_weight_set <- dataset_list[[2]]
  ##

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

```{r Save target encoding weights, include=FALSE, eval=FALSE}
save(target_encoding_weights,file='rda/target_encoding_weights.rda')
```

```{r Load target encoding weights, include=FALSE}
# required only to generate report or resume modeling quickly
load('rda/target_encoding_weights.rda')
```

penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
Similarly delay in payment approval is used to incorporate role of customer zip code in delaying this processing, which maybe due to fraud analysis etc.
```{r Prepare training and test data for model tuning and development,include=FALSE}
dataset_list <- prepare_data_for_training_testing(train_subset,test_subset,FALSE)
train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]

# cleanup
train_subset <- train_subset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
test_subset <- test_subset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
```

@@

```{r knn model data scaled under test, include=FALSE, eval=FALSE}
## TODO make echo=FALSE 
set.seed(seed=1,sample.kind = 'Rounding')
preProcessScalingValues <- preProcess(train_subset %>% select(-number_of_hours), method = c("center", "scale"))

@@

tuningControl=trainControl(method = "cv", number = 5)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)

@@

performance_metric <- performance_metric %>% 
  add_row(type='Knn model(partial dataset)',
          value=RMSE((scaled_test_subset %>% pull(number_of_hours)),predictions))

performance_metric
#183.7767 for k=63

@@

```{r rf model, include= FALSE, eval = FALSE}
# TODO make it echo=FALSE
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 4)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5)),trControl=tuningControl,ntree=200)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

performance_metric <- performance_metric %>% 
  add_row(type='Random forest model(partial dataset)',
          value=RMSE((test_subset %>% pull(number_of_hours)),predictions))

performance_metric
# 174.2496 for mtry 2. Technically there are total of 11 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.

@@

```{r Save performance metric, include=FALSE, eval=FALSE}
save(performance_metric,file='rda/performance_metric.rda')
```

```{r Load performance metric, include=FALSE}
# required only to generate report
load('rda/performance_metric.rda')
```

```{r ensemble, include=FALSE, eval= FALSE}
load('rda/knn_model.rda')
load('rda/rf_model.rda')

predictions_knn <- predict(knn_fit,scaled_test_subset %>% select(-number_of_hours))
predictions_rf <- predict(rf_fit,test_subset %>% select(-number_of_hours))

predictions_df <- data.frame(predictions_knn,predictions_rf)

RMSE(rowMeans(predictions_df),(test_subset %>% pull(number_of_hours) ))
176.5041
```

```{r Prepare dataset for final model training, include = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')

validation_dataset <- validation_dataset %>% 
  compute_volume_using_product_info() %>% 
  extract_distance_between_seller_customer() %>% 
  compute_rough_customer_location()

dataset_list <- prepare_data_for_training_testing(training_dataset,validation_dataset,FALSE)
training_dataset <- dataset_list[[1]]
validation_dataset <- dataset_list[[2]]

# cleanup
training_dataset <- training_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
validation_dataset <- validation_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
```

```{r Final Model Training, include=FALSE,eval = FALSE}
# TODO marked eval=FALSE to prevent it from running during report generation
set.seed(seed=1,sample.kind = 'Rounding')

mtry_to_use <- rf_fit$finalModel$mtry
ntree_to_use <- 500 #rf_fit$finalModel$ntree

final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)

final_model

save(final_model,file='rda/final_model.rda')
```

```{r Final model validation, include = FALSE, eval=FALSE}
# TODO marked eval=FALSE to prevent it from running during report generation
predictions <- predict(final_model,validation_dataset %>% select(-number_of_hours))

performance_metric <- performance_metric %>% 
  add_row(type='Final model',
          value=RMSE((validation_dataset %>% pull(number_of_hours) ),predictions))
# 189.2705 for mtry=2, ntree=500

performance_metric
```
